Long name of the paper

### 2018-09-08 anzjs2018/sections/03_models.tex
model writing

>>>>

The final model $\bM_3$ demonstrated here will allow acceleration to vary.
\begin{align*}
\label{eq:model_m3}
x_{k|k-1} &= x(t_k) \\
\intertext{where states are recursively calculated back to $t_{k-1}$}
x(t) &= x(t - 1) + \dot x(t) \\
\dot x(t) &= \dot x(t-1) + \ddot x(t) \\
\ddot x(t) &= \ddot x(t-1) + w_t,\quad
w_t \sim \mathcal{N}(0,\sigma_x^2)
\end{align*}
By truncating the acceleration, 
the speed of the vehicle can be maintained between 0 and 30~m/s.
Additionally, when approaching a bus stop that will be stopped at,
the acceleration will be trunctated to below zero,
however no physical deceleration will be modeled as this is too complex
and stopping behaviour of buses can be sudden 
(for example when passengers signal late that they wish to board or disembark).
After servicing a stop, the bus will need to accelerate up to speed again.

\subsubsection{Updating state using the observation likelihood}
\label{sec:pf_update}

The update step involves comparing the predicted state $\bX_{k|k-1}$
with the observed state $\bY_k$,
using the measurement function $h$ to convert between the two state spaces,
and a likelihood function.
The measurement function is a deterministic function that uses only the 
distance traveled, $x_k$, along a known path $\bR$ to obtain a GPS coordinate.
The inverse measurement function, which is used in other approaches,
is \emph{not deterministic}, and instead uses an optimisation to
find the ``estimated'' distance traveled,
which invariably leads to issues in the implementation.
By using a particle filter approach, we avoid using $h^{-1}$
and instead determine the liklihood using $\bY_k$ directly.

For the likelihood, the GPS error is assumed symmetrical about the true
position of the vehicle, with a radius of $\sigma_y^2$ meters.
Define a projection function $g(\cdot | \bY)$ that projects points
onto a cartesian surface (``Flat Earth''),
so that the distance between points is the real Earth distance.
Then the distance between the points is
\begin{align*}
\delta_k &= \sqrt{\left(g(h(X_k))_1 - g(\bY_k)_1\right)^2 +
    \left(g(h(X_k))_2 - g(\bY_k)_2\right)^2} \\
\intertext{which can be standardised by the GPS error}
\tilde \delta_k &=
\sqrt{\left(\frac{g(h(X_k))_1 - g(\bY_k)_1}{\sigma_y}\right)^2 +
    \left(\frac{g(h(X_k))_2 - g(\bY_k)_2}{\sigma_y}\right)^2} =\frac{\delta_k}{\sigma_y} \\
\intertext{Since we assume symmetric variance, this is actually}
\tilde\delta_k &= \sqrt{Z_1^2 + Z_2^2}, \quad
Z_1,Z_2 \sim \mathcal{N}(0,1)
\end{align*}
The sum of two normal random variables has an Exponential distribution with rate $\frac{1}{2}$,
therefore 
\begin{equation}
\label{eq:lhood_exp}
\left(\frac{\delta_k}{\sigma_y}\right)^2 \sim \mathrm{Exp}\left(\frac{1}{2}\right)
\end{equation}

Combining this together, we use a distance function $d(\bY_1, \bY_2)$ that calculates
the geographic distance between two GPS coordinates,
the likelihood is
\begin{equation}
\label{eq:lhood}
f(\bY_k | \hat\bX_{k|k-1}, \sigma_y^2) =
\frac{1}{2} \exp\left\{-\frac{d(\bY_k, h(\hat\bX_{k|k-1}))^2}{2\sigma^2}\right\}
\end{equation}
